More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-02-07 2016-Weak-FH-Bisim/WeakBisim.tex
Reu RAB

In the following definitions, we extensively use indexed structures
(maps) over some countable indexed sets. The indices will usually be
integers, bounded or not. Such an indexed family is denoted as
follows: $a_i^{i\in I}$ is a family of elements $a_i$ indexed over the
set $I$. Such a family
is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
To specify the set over which the structure is indexed, 
indexed structures are always denoted with an exponent of the form $i\in I$
(arithmetic only appears in the indexes if necessary).
Consequently, $a_i^{i\in I}$ defines first $I$ the set over which the
family is indexed, and then $a_i$ the elements of the family.

For example $a^{i\in\{3\}}$ is
the mapping with a single entry $a$ at index $3$; exceptionally, such mappings with
only a few entries will also be denoted $(3\mapsto a)$. \TODO{We need to define update as $A \uplus \{j \mapsto a\}$ or $A[j \mapsto a]$}
When this is not ambiguous, we shall use abusive vocabulary and
notations for sets, and typically write ``indexed set over I'' when  
formally we should speak of multisets, and ``$x\in
A_i^{i\in I}$'' to mean $\exists i\in I.\, x=A_i$.
An empty family is denoted $[]$. % (it can be defined as $a_i^{i\in\emptyset}$).
To simplify equations, an indexed set can be denoted $\set{M}$
instead of $M_l^{l\in L}$ when $L$ is irrelevant.

Let $\uplus$ be the disjoint union operator on indexed sets. 
The elements
of the union are then accessed by using an index of one of the two
joined families.
We suppose here that disjoint unions are always well defined; this
requires to choose disjoint  
indices in the definition of the component system (e.g. name of
methods of different interfaces).
We denote $y\gets x$ a substitution. The application of the substitution is denoted
$\subst{y\gets x}$, the operation replaces in a term all occurrences 
of the term $y$ by the term $x$ (note that the absence of binders makes this operation 
trivial).

@@

\ERIC{Keep minimum defs here too be independent from FORTE}

@@

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\cup{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined inductively as:
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\! \{ \mylangle S,s_0, \to\myrangle \}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \bigcup_{i\in 
I}\Leaves(\pNet_i)}
\end{array}\]
\end{itemize}
\end{definition}

@@

  For any pNet p, let $\Leaves(p) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

@@

\RAB{Exemple  Fig3 a revoir: Les regles ne passent... }

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNets $\pNet'$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNet'\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNet'\},\Sort_j^{j\in J\setminus \{j_0\}},
	\overline{\symb{SV}}\myrangle
	\]
\end{definition}
\TODO{$\uplus\{j_0\mapsto \pNet'\}$ non defini}

@@

Two pNets are Weak FH-bisimilar if there exist a relation between their associated open
automata that is an Weak FH-bisimulation and their initial states are equivalent.
\TODO{define equivalent}
 \end{definition}

\TODO{Here we should include:
  \\  - A definition for Weak FH-Bisimulation
  \\  - A sufficient condition for an open PNet to preserve Weak-FH,
  including: 1) no non-deterministic Tau-choice in the pLTSs, 2) a
  synch vector propagating Tau from a subnet $pNet_i$ mandatory
  everytime $pNet_i$ can do some visible action in the current
  context, 3) if we have a vector <Tau, -, a, - > --> b with ``a'' a visible
  action of some CLOSED subnet, then b=Tau. 
  \\  - the corresponding Congruence theorem
  \\- An example
}

\TODO{Eric: ask Elie, Hubert or others about interesting weak
  equationnal theories for Lotos -- DONE, no answer..., so I got some examples from \cite{Lotos:ISO}}

@@

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation $\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
 t=t'\uplus s'' \land s''\in \mathcal{S}_P \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ \TODO{$\uplus$ entre states?} 
 where $\mathcal{S}_S$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{i}^{i \in L_P\uplus L}},\ostate{{t}_{i}^{i \in L'}\uplus 
 	{s}_{i}^{i \in L_P}}|\Pred_{s,t})\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
 {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
 	%States(I_1)
 	\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{i}^{i \in L_P\uplus L}}\OTarrow {v} \ostate{{s'}_{i}^{~i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L},
 		\{\xrightarrow{b_j}_j\}^{j\in J'}, 
 		\Pred\,',  
 		\Post\,'}
 	{\ostate{s_{i}^{i\in L_P}} \OTarrow {v}
 		\ostate{s_{i}'^{\,i\in L_P}}}
 }\\
 Q\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L},
 		\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post\,''}
 	{\ostate{s_{i}^{i\in L}} \OTarrow {v_Q}
 		\ostate{s_{i}'^{\,i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,'}
 \land {\Pred\,}'')$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{i}^{i \in 
 		L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\inferrule*[myfraction=\reddottedrule]
 {\{t_{i}\xrightarrow{a_{i x}}_i t_{i x}\}^{i\in I_{x}},
 	\{\xrightarrow{b_{j x}}_j\}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{t_{i}^{i\in L'}} \OTarrow {v_x} \ostate{t_{i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
 		L'}}|\Pred_{s x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred\,''
 \Rightarrow \bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 \Pred_{OT_x}
 \land v_Q\!=\!v_x \land  
 \Pred_{s x}\subst{\Post\,''}\subst{\Post_{OT_x}}\right)$

>>>>

In the following definitions, we extensively use indexed structures
(maps) over some countable indexed sets. The indices will usually be
integers, bounded or not. Such an indexed family is denoted as
follows: $a_i^{i\in I}$ is a family of elements $a_i$ indexed over the
set $I$. Such a family
is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$, and we will also use mapping 
notations to manipulate indexed sets.
To specify the set over which the structure is indexed, 
indexed structures are always denoted with an exponent of the form $i\in I$
(arithmetic only appears in the indexes if necessary).
Consequently, $a_i^{i\in I}$ defines first $I$ the set over which the
family is indexed, and then $a_i$ the elements of the family.

For example $a^{i\in\{3\}}$ is
the mapping with a single entry $a$ at index $3$; exceptionally, such mappings with
only a few entries will also be denoted $(3\mapsto a)$.
%The operation  $A[j \mapsto a]$ updates the value associated to $j$ in map $A$ so that 
%it 
%now corresponds to $a$. 
When this is not ambiguous, we shall use abusive vocabulary and
notations for sets, and typically write ``indexed set over I'' when  
formally we should speak of multisets, and ``$x\in
A_i^{i\in I}$'' to mean $\exists i\in I.\, x=A_i$.
An empty family is denoted $[]$. % (it can be defined as $a_i^{i\in\emptyset}$).
To simplify equations, an indexed set can be denoted $\set{M}$
instead of $M_l^{l\in L}$ when $L$ is irrelevant.

$\uplus$ is the disjoint operator on sets. We extend it to  disjoint union  of indexed 
sets defined by the merge of the 
two sets provided they are indexed on disjoint families.
The elements
of the union of two indexed sets are then accessed by using an index of one of the two
joined families.
We suppose here that disjoint unions are always well defined; this
requires to choose disjoint  
indices in the definition of the component system (e.g. name of
methods of different interfaces).
We denote $y\gets x$ a substitution. The application of the substitution is denoted
$\subst{y\gets x}$, the operation replaces in a term all occurrences 
of the term $y$ by the term $x$ (note that the absence of binders makes this operation 
trivial).

@@

@@

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\TODO{Not very nice but at least of the right kind for the definition of composition}
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i=\mylangle S,s_0, \to\myrangle\}}
\end{array}\]
\end{itemize}
\end{definition}

@@

  For any pNet p, let $\Leaves(p) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
\TODO{this definition is not valid because L is undefined and leaves only exist for pNets 
-> REMOVE? this would mean that a pNet is not a pLTS ...}
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

@@

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNets $\pNet'$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNet'\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNet'\},\Sort_j^{j\in J\setminus \{j_0\}},
	\overline{\symb{SV}}\myrangle
	\]
\end{definition}

@@

Two pNets are Weak FH-bisimilar if there exist a relation between their associated open
automata that is an Weak FH-bisimulation and their initial states are equivalent.
\TODO{LUDO: define equivalent}
 \end{definition}

\TODO{Here we should include:
  \\  - A sufficient condition for an open PNet to preserve Weak-FH,
  including: 1) no non-deterministic Tau-choice in the pLTSs, 2) a
  synch vector propagating Tau from a subnet $pNet_i$ mandatory
  everytime $pNet_i$ can do some visible action in the current
  context, 3) if we have a vector <Tau, -, a, - > --> b with ``a'' a visible
  action of some CLOSED subnet, then b=Tau. 
  \\  - the corresponding Congruence theorem
  \\- An example
}

@@

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation\footnote{$\uplus$ is defined on sets as the disjoint union 
 of the contained indexed 
 sets} $\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
 t=t'\uplus s'' \land s''\in \mathcal{S}_P \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$

 where $\mathcal{S}_S$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{i}^{i \in L_P\uplus L}},\ostate{{t}_{i}^{i \in L'}\uplus 
 	{s}_{i}^{i \in L_P}}|\Pred_{s,t})\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
 {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
 	%States(I_1)
 	\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{i}^{i \in L_P\uplus L}}\OTarrow {v} \ostate{{s'}_{i}^{~i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L},
 		\{\xrightarrow{b_j}_j\}^{j\in J'}, 
 		\Pred\,',  
 		\Post\,'}
 	{\ostate{s_{i}^{i\in L_P}} \OTarrow {v}
 		\ostate{s_{i}'^{\,i\in L_P}}}
 }\\
 Q\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L},
 		\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post\,''}
 	{\ostate{s_{i}^{i\in L}} \OTarrow {v_Q}
 		\ostate{s_{i}'^{\,i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,'}
 \land {\Pred\,}'')$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{i}^{i \in 
 		L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\inferrule*[myfraction=\reddottedrule]
 {\{t_{i}\xrightarrow{a_{i x}}_i t_{i x}\}^{i\in I_{x}},
 	\{\xrightarrow{b_{j x}}_j\}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{t_{i}^{i\in L'}} \OTarrow {v_x} \ostate{t_{i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
 		L'}}|\Pred_{s x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred\,''
 \Rightarrow \bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 \Pred_{OT_x}
 \land v_Q\!=\!v_x \land  
 \Pred_{s x}\subst{\Post\,''}\subst{\Post_{OT_x}}\right)$
